Remote Collection of Respiratory Samples (NACFC 2023)

### 2023-10-29 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(irr)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("1", "5"), labels = c("American Indian or Alaska Native", "White"))
  )
# Shorten organism names
bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
bugs = bugs[1:16]
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
bug_names=c("Methicillin sensitive Staph aureus", 
            "Methicillin resistant Staph aureus","A. xylosoxidans", 
            "Achromobacter (not xylosoxidans)","Acinetobacter","H. influenzae", 
            "P. aeruginosa", "P. aeruginosa (mucoid)", 
            "Pseudomonas species (not aeruginosa)", "B. cepacia complex",
            "Stenotrophomonas", "Aspergillus", "Penicillium species", 
            "Sceosporium species", "M. avium complex", "M. abscessus" )
```

@@

```{r}
fleiss = function(bug){
  d = df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))]
  colnames(d) = c("Clinic","Delayed","Home")
  d[d=="Checked"] = "Present"
  d[d=="Unchecked"] = "Absent"
  fk = kappam.fleiss(d)
  cat("\n")
    cat("##", bug_names[which(bugs==bug)])
    cat("\n")
    print(ftable(d))
    cat("\n")
    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
    cat("\n")
}
```

@@

```{r message=FALSE}
f = df %>% 
  select(Subject.ID,contains("Methicillin.sensitive"),contains("H..influenzae."))
colnames(f) = c("ID",paste("MSSA",c("Clinic","Delayed","Home")),
                paste("H. influenzae",c("Clinic","Delayed","Home")))

counts=f %>% 
  pivot_longer(`MSSA Clinic`:`H. influenzae Home`,
                       names_to = c("Bug","Test"),
                       names_pattern = "(.*) (.*)") %>%
  filter(value=="Checked") %>%
  group_by(Bug,ID) %>%
  summarise(Test=paste0(Test,collapse = "+")) %>%
  count(Test) %>%
  mutate(prop = prop.table(n))

ggplot(counts,aes(x=Bug,y=n,fill=Test,label=n))+
  geom_col(colour = "black", position = "fill") +
  geom_text(aes(label = n), position = position_fill(vjust = 0.5)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Pastel1")+
  xlab("Pathogen")+ylab("% of Samples")
```

@@

```{r results='asis'}
# For those who tested negative for any bugs in clinic, did they also test
# negative for all at home
f = df %>% select(Subject.ID,contains (bugs)) %>% 
  mutate(across(-Subject.ID,
                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1)) %>%
  rowwise() %>%
  mutate(`Clinic % Positive` = mean(c_across(all_of(bugs)))*100,
         `Delayed % Positive` = mean(c_across(all_of(paste0(bugs,".1"))))*100,
         `Home % Positive` = mean(c_across(all_of(paste0(bugs,".2"))))*100) %>%
  select(Subject.ID,`Clinic % Positive`:`Home % Positive`) %>%
  filter(`Clinic % Positive`==0)
knitr::kable(f)
```

```{r}
f = f %>%
  mutate(across(c("Clinic % Positive","Delayed % Positive","Home % Positive"),~.x==0))
kappam.fleiss(f[,2:4])
```

>>>>

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(irr)
library(gtsummary)
df <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical_2023-06-20.csv", na.strings = "")
surveys <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys_2023-06-20.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("1", "5"), labels = c("American Indian or Alaska Native", "White"))
  )
# Shorten organism names
bugs <- colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.", colnames(df))]
bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
bugs <- bugs[1:16]
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
bug_names <- c(
  "Methicillin sensitive Staph aureus",
  "Methicillin resistant Staph aureus", "A. xylosoxidans",
  "Achromobacter (not xylosoxidans)", "Acinetobacter", "H. influenzae",
  "P. aeruginosa", "P. aeruginosa (mucoid)",
  "Pseudomonas species (not aeruginosa)", "B. cepacia complex",
  "Stenotrophomonas", "Aspergillus", "Penicillium species",
  "Sceosporium species", "M. avium complex", "M. abscessus"
)
```

@@

```{r}
fleiss <- function(bug) {
  d <- df[, c(bug, paste0(bug, ".1"), paste0(bug, ".2"))]
  colnames(d) <- c("Clinic", "Delayed", "Home")
  d[d == "Checked"] <- "Present"
  d[d == "Unchecked"] <- "Absent"
  fk <- kappam.fleiss(d)
  cat("\n")
  cat("##", bug_names[which(bugs == bug)])
  cat("\n")
  print(ftable(d))
  cat("\n")
  cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value, digits = 3, eps = 0.001), ")"))
  cat("\n")
}
```

@@

```{r message=FALSE}
f <- df %>%
  select(Subject.ID, contains("Methicillin.sensitive"), contains("H..influenzae."))
colnames(f) <- c(
  "ID", paste("MSSA", c("Clinic", "Delayed", "Home")),
  paste("H. influenzae", c("Clinic", "Delayed", "Home"))
)

counts <- f %>%
  pivot_longer(`MSSA Clinic`:`H. influenzae Home`,
    names_to = c("Bug", "Test"),
    names_pattern = "(.*) (.*)"
  ) %>%
  filter(value == "Checked") %>%
  group_by(Bug, ID) %>%
  summarise(Test = paste0(Test, collapse = "+")) %>%
  count(Test) %>%
  mutate(prop = prop.table(n))

ggplot(counts, aes(x = Bug, y = n, fill = Test, label = n)) +
  geom_col(colour = "black", position = "fill") +
  geom_text(aes(label = n), position = position_fill(vjust = 0.5),size=5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Pastel1") +
  xlab("Pathogen") +
  ylab("% of Samples") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

@@

```{r results='asis'}
# For those who tested negative for any bugs in clinic, did they also test
# negative for all at home
f <- df %>%
  select(Subject.ID, contains(bugs)) %>%
  mutate(across(
    -Subject.ID,
    ~ as.numeric(factor(.x, levels = c("Unchecked", "Checked"))) - 1
  )) %>%
  rowwise() %>%
  mutate(
    `Clinic % Positive` = mean(c_across(all_of(bugs))) * 100,
    `Delayed % Positive` = mean(c_across(all_of(paste0(bugs, ".1")))) * 100,
    `Home % Positive` = mean(c_across(all_of(paste0(bugs, ".2")))) * 100
  ) %>%
  select(Subject.ID, `Clinic % Positive`:`Home % Positive`) %>%
  filter(`Clinic % Positive` == 0)
knitr::kable(f)
```

```{r}
f <- f %>%
  mutate(across(c("Clinic % Positive", "Delayed % Positive", "Home % Positive"), ~ .x == 0))
kappam.fleiss(f[, 2:4])
```
